EWS analysis

### 2021-04-14 various_stuff/EWS functions/EWS analyses.Rmd
reorganisation of vignettes
now need to get them working!

---
title: "Diversity and the nature of responses to environmental change."
author: "Owen Petchey & Jennifer Probst"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 4
    number_sections: true
    code_folding: hide
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Setup R

```{r setup}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

library(tidyverse)
library(deSolve)
library(rootSolve)

set.seed(13)

run_sims <- FALSE
```

# Introduction

The aims here are:

* Reproduce the results of [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x).
* Add an investigation of how changes in tolerances alter the response to environmental change.

We also add:

* Option to include random noise in substrate concentrations.
* Option to temporally vary oxygen diffusivity.
* Option to include immigration.
* Option to include a minimum abundances.

Here we use only the published (Bush et al 2017) dynamical model, in which there is not biodiversity within each of the three functional groups. That is, there is only one parameter set per functional group.

Here is the original publication:
https://www.nature.com/articles/s41467-017-00912-x

And here is the supplementary information, including the table of parameter values:
https://static-content.springer.com/esm/art%3A10.1038%2Fs41467-017-00912-x/MediaObjects/41467_2017_912_MOESM1_ESM.pdf

Note that we here use the notation from the table S1 of the supplementary information, and not the notation in the main text. (There is only one small difference between these: In the main text the notation is simplified to y_S_PB (which is y_SR_PB ($y_{SR,PB}$) in Table S1) and y_S_SB (which is y_SO_SB ($y_{SO,SB}$) in Table S1).

# Methods

## Model dimensions

* Time: hours
* Volume: litres
* Substrate quantity: micromoles
* Organism quantity: cells

## Dynamic model

```{r}
source("functions/bushplus_dynamic_model.r")
```

## Defaults

The framework used for running a simulation, and for varying aspects of a simulation run is:

* Define the default simulation conditions.
* Use these defaults unless other conditions are specified.
* The function call `run_simulation()` would therefore run a simulation with default conditions.
* Whereas the function call `run_simulation(initial_state = init_state_favouring_anoxic)` would run the simulation with specified initial state values.

### Default parameter values

```{r}
default_parameter_values <- c(

  ## maximum specific growth rates
  g_max_CB = 0.05,
  g_max_PB = 0.07,
  g_max_SB = 0.1,

  ## half-saturation constants
  k_PB_SR = 10,
  k_SB_SO = 5,
  k_CB_P = 0.2,
  k_PB_P = 0.5,
  k_SB_P = 0.5,

  ## half-inhibition constant
  h_SR_CB = 300,
  h_O_PB = 100,
  h_O_SB = 100,

  ## yield (cells per micromole)
  Y_SO_SB = 3.33e7,  
  Y_SR_PB = 1.25e7, ## Y_S_PB in the main text
  y_P_CB = 1.67e8,
  y_P_PB = 1.67e8,
  y_P_SB = 1.67e8,

  ## oxygen production per microbial cell
  P_CB = 6e-9,

  ## mortality rate
  m_CB = 0.02,
  m_PB = 0.028,
  m_SB = 0.04,

  ## substrate diffusivity
  a_S = 0.001,
  a_O = 8e-4,
  a_P = 0.01,

  ## background substrate concentration
  back_SR = 300,
  back_SO = 300,
  back_O = 300,
  back_P = 9.5,

  ## immigration rates (cells per time)
  i_CB <- 0,
  i_PB <- 0,
  i_SB <- 0,

  ## oxidisation rate of reduced sulphur
  c = 4e-5
  )
```

### Default event

#### Default event definition

Used to add noise and to ensure minimum abundance. Noise is added to the resource concentrations, and not to the organism abundances. Minimum organism abundance is set. I.e. organismal abundance cannot go below some very low level.

```{r}
source("functions/default_event_definition.r")
```

#### Default event schedule

This is when an event happens

```{r}
default_event_interval <- 10
```

#### Default noise event options

Default to no noise:

```{r}
default_noise_sigma <- 0
```

#### Default minimum abundance event

Default to following minimums

```{r}
default_minimum_abundances <- rep(1e2, 3)
names(default_minimum_abundances) <- c("CB", "PB", "SB") 
```

### Default simulation duration

```{r}
default_sim_duration <- 2000
default_sim_sample_interval <- 1
```

### Default oxygen diffusivity (a) forcing

Here we specify no forcing of a_0 (used unless set elsewhere).

```{r}
default_log10a_series <- c(log10(default_parameter_values["a_O"]),
                           log10(default_parameter_values["a_O"]))
```

### Default initial state

Initial conditions favouring anoxic.

```{r}
default_initial_state <- c(N_CB = 5e1,
           N_PB = 1e7,
           N_SB = 1e7,
           SO = 300,
           SR = 300,
           O = 1e1,
           P = 1e1)
```

## Simulation function

```{r}
source("functions/run_simulation.r")
```

## Helper functions

```{r}
source("functions/plot_dynamics.r")
source("functions/ss_expt_a_N.r")
source("functions/get_nonlinearity.r")
source("functions/get_hysteresis_range.r")
```

## Code test

```{r}
# dynamic_model = bushplus_dynamic_model
# event_definition = default_event_definition
# parameter_values = default_parameter_values
# event_interval = default_event_interval
# noise_sigma = default_noise_sigma
# minimum_abundances = default_minimum_abundances
# sim_duration = default_sim_duration
# sim_sample_interval = default_sim_sample_interval
# log10a_forcing = default_log10a_forcing
# initial_state = default_initial_state
# solver_method = "radau"
# 
# 
# times <- seq(0,
#              sim_duration,
#              by = sim_sample_interval)
# 
# event_times <- seq(min(times),
#                    max(times),
#                    by = event_interval)  
# 
# log10a_forcing <- approxfun(x = log10a_forcing[,1],
#                             y = log10a_forcing[,2],
#                             method = "linear",
#                             rule = 2)
# assign("log10a_forcing", log10a_forcing, envir = .GlobalEnv)
# 
# out <- as.data.frame(ode(y = initial_state,
#                          times = times,
#                          func = dynamic_model,
#                          parms = parameter_values,
#                          method = solver_method,
#                          events = list(func=event_definition,
#                                        time=event_times)))
# 
# sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
#                            event_definition = default_event_definition,
#                            parameter_values = default_parameter_values,
#                            event_interval = default_event_interval,
#                            noise_sigma = default_noise_sigma,
#                            minimum_abundances = default_minimum_abundances,
#                            sim_duration = default_sim_duration,
#                            sim_sample_interval = default_sim_sample_interval,
#                            log10a_forcing = default_log10a_forcing,
#                            initial_state = default_initial_state,
#                            solver_method = "radau")

#sim_res <- run_simulation()

#plot_dynamics(sim_res)

# Results: Reproducing previous results

## Initial conditions favouring anoxic, no noise, no forcing

Figure 2a, b in Bush et al.

```{r}

init_state_favouring_anoxic <- c(N_CB = 5e1,
           N_PB = 1e7,
           N_SB = 1e7,
           SO = 300,
           SR = 300,
           O = 1e1,
           P = 1e1)

sim_res <- run_simulation(initial_state = init_state_favouring_anoxic)
plot_dynamics(simulation_result = sim_res)

#ggsave('bacteria_no_noise.png')
```

Population dynamics seem as in the paper.
Resource dynamics fluctuate more than in the paper.

## Initial conditions favouring anoxic, with noise, no forcing

Figure 2a, b in Bush et al, but with noise

```{r}
## add some noise
noise_sigma_0.01 <- 0.01
sim_res <- run_simulation(initial_state = init_state_favouring_anoxic,
                          noise_sigma = noise_sigma_0.01)
plot_dynamics(sim_res)
#ggsave("high_SB_PB.png")
```

## Initial conditions favouring oxic, no noise, no forcing

Figure 2c, d in Bush et al.

```{r} 
init_state_favouring_oxic <- c(N_CB = 1e7,
           N_PB = 1e5,
           N_SB = 1e5,
           SO = 500,
           SR = 50,
           O = 30,
           P = 10)

sim_res <- run_simulation(initial_state = init_state_favouring_oxic)

plot_dynamics(sim_res)
##ggsave("high_CB.png")
```
Population and resource dynamics seen as those in the paper.

## Initial conditions favouring oxic, with noise, no forcing

Figure 2c, d in Bush et al. but with noise.

```{r}

sim_res <- run_simulation(initial_state = init_state_favouring_oxic,
                          noise_sigma = noise_sigma_0.01)

plot_dynamics(sim_res)
##ggsave("high_CB.png")
```

# Results: Composition/diversity and nature of response

```{r}
ssfind_minimum_abundances <- rep(0, 3)
names(ssfind_minimum_abundances) <- c("CB", "PB", "SB") 
ssfind_simulation_duration <- 10000
ssfind_simulation_sampling_interval <- ssfind_simulation_duration
ssfind_event_interval <- ssfind_simulation_duration
ssfind_parameters <- default_parameter_values
# ssfind_init_state <- c(N_CB = NA, 
#                        N_PB = NA,
#                        N_SB = NA,
#                        SO = 500,
#                        SR = 50,
#                        O = 30,
#                        P = 10)
##ssfind_init_state <- default_initial_state
ssfind_init_state <- init_state_favouring_anoxic
ssfind_init_state <- c(N_CB = NA, #this one is varied
           N_PB = 1e8,
           N_SB = 1e8,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)
grid_num_a <- 100
a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
grid_num_N <- 2
```

### Each composition

#### No organisms

```{r eval = run_sims}

initial_CBs <- 0
initial_PBs <- 0
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_no_biology <- ss_by_a_N(expt)
saveRDS(ss_no_biology, "data/ss_no_biology.RDS")
```

```{r}
ss_no_biology <- readRDS("data/ss_no_biology.RDS")
```

```{r}
ss_no_biology %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_no_biology %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### CB

```{r eval = run_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 0
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_only_CB <- ss_by_a_N(expt)
saveRDS(ss_only_CB, "data/ss_only_CB.RDS")
```

```{r}
ss_only_CB <- readRDS("data/ss_only_CB.RDS")
```

```{r}
tempx1 <- ss_only_CB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.))
tempx1 %>%
  ggplot(aes(x = a, y = log10(Quantity), col = Species)) +
  geom_path(size=0.8) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  theme_light() +
  ylim(0,10) +
  geom_path(data = filter(tempx1, Species == "PB"),
            aes(y=0.2), size = 0.8) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="Only CB")

ggsave("figures/CB.png", width = 3, height = 2)
```

```{r}
ss_only_CB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### SB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 0
initial_SBs <- 10^seq(0, 8, length=grid_num_N)
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_only_SB <- ss_by_a_N(expt)

saveRDS(ss_only_SB, "data/ss_only_SB.RDS")
```

```{r}
ss_only_SB <- readRDS("data/ss_only_SB.RDS")
```

```{r}
ss_only_SB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### PB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 10^seq(0, 8, length=grid_num_N)
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_only_PB <- ss_by_a_N(expt)

saveRDS(ss_only_PB, "data/ss_only_PB.RDS")
```

```{r}
ss_only_PB <- readRDS("data/ss_only_PB.RDS")
```

```{r}
tempx1 <- ss_only_PB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.))
tempx1 %>%
  ggplot(aes(x = a, y = log10(Quantity), col = Species)) +
  geom_path(size = 0.8) +
  theme_light() +
  ylim(0,10) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  geom_path(data = filter(tempx1, Species == "CB"),
            aes(y=0.2), size = 0.8)+
  geom_path(data = filter(tempx1, Species == "PB"),
            size = 0.8) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="Only PB")

ggsave("figures/PB.png", width = 3, height = 2)

```{r}
ss_only_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### SB and PB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 10^seq(0, 8, length=grid_num_N)
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)
expt$N_SB <- expt$N_PB

ss_only_SB_PB <- ss_by_a_N(expt)
saveRDS(ss_only_SB_PB, "data/ss_only_SB_PB.RDS")
```

```{r}
ss_only_SB_PB <- readRDS("data/ss_only_SB_PB.RDS")
```

```{r}
ss_only_SB_PB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### SB and CB

```{r eval = run_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 0
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)
expt$N_SB <- expt$N_CB

ss_only_SB_CB <- ss_by_a_N(expt)
saveRDS(ss_only_SB_CB, "data/ss_only_SB_CB.RDS")
```

```{r}
ss_only_SB_CB <- readRDS("data/ss_only_SB_CB.RDS")
```

```{r}
ss_only_SB_CB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB_CB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:5) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### CB and PB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 10^seq(0, 8, length=grid_num_N)
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)
expt$N_CB <- expt$N_PB

ss_only_CB_PB <- ss_by_a_N(expt)
saveRDS(ss_only_CB_PB, "data/ss_only_CB_PB.RDS")
```

```{r}
ss_only_CB_PB <- readRDS("data/ss_only_CB_PB.RDS")
```

```{r}
ss_only_CB_PB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### CB, SB, and PB

```{r eval = run_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 1e5
initial_SBs <- 1e5
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_CB_SB_PB <- ss_by_a_N(expt)
saveRDS(ss_CB_SB_PB, "data/ss_CB_SB_PB.RDS")
```

```{r}
ss_CB_SB_PB <- readRDS("data/ss_CB_SB_PB.RDS")
```

```{r}
ss_CB_SB_PB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species),
            size=0.8) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  theme_light() +
  ylim(0,10) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="CB, SB, & PB") 
ggsave("figures/CB_SB_PB.png", width = 3, height = 2)

```{r}
ss_CB_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

### All response measures

```{r eval = run_sims}
all_ss <- bind_cols(readRDS("data/ss_no_biology.RDS"),
                    composition = rep("No organisms", grid_num_a*grid_num_N),
                    direction = c(rep("up", grid_num_a),
                                  rep("down", grid_num_a)))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_CB.RDS"),
                              composition = rep("CB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_SB.RDS"),
                              composition = rep("SB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_PB.RDS"),
                              composition = rep("PB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_SB_PB.RDS"),
                              composition = rep("SB-PB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_SB_CB.RDS"),
                              composition = rep("SB-CB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_CB_PB.RDS"),
                              composition = rep("CB-PB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_CB_SB_PB.RDS"),
                              composition = rep("CB-SB-PB",
                                                grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))

x_long <- all_ss %>%
  gather(key = Species, value = Quantity, 2:8)
x_wide <- x_long %>%
  select(-initial_N_CB) %>%
  spread(key = direction, value=Quantity, drop=T)

resp_summary <- x_wide %>%
  group_by(composition, Species) %>%
  summarise(hyst_1 = mean(abs(log10(up) - log10(down))),
            hyst_2 = get_hysteresis_range(up, down, a),
            nl_up = get_nonlinearity(a, log10(up)),
            nl_down = get_nonlinearity(a, log10(down))) %>%
  ungroup() %>%
  gather(key = resp_measure, value = value, 3:6) %>%
  mutate(composition = fct_relevel(composition,
                                   "No organisms",
                                   "CB",
                                   "SB",
                                   "PB",
                                   "CB-PB",
                                   "SB-CB",
                                   "SB-PB",
                                   "CB-SB-PB"),
         num_func_groups = case_when(composition =="No organisms" ~ 0,
                                     composition == "CB" ~ 1,
                                     composition == "SB" ~ 1,
                                     composition == "PB" ~ 1,
                                     composition == "CB-PB" ~ 2,
                                     composition == "SB-CB" ~ 2,
                                     composition == "SB-PB" ~ 2,
                                     composition == "CB-SB-PB" ~ 3
                                     ),
         var_type = ifelse(str_sub(Species, 1, 1)=="N", "Organism", "Substrate")
         )
saveRDS(resp_summary, "data/resp_summary.RDS")
```

```{r}
resp_summary <- readRDS("data/resp_summary.RDS")
resp_summary %>%
  ggplot() +
  geom_point(aes(x = composition, y = value)) +
  facet_grid(Species ~ resp_measure, scales = "free") +
  coord_flip()

### Nonlinearity

#### All species nonlinearity

```{r}
resp_summ1 <- resp_summary %>%
  filter(resp_measure %in% c("nl_down", "nl_up")) %>%
  group_by(composition, Species, num_func_groups, var_type) %>%
  summarise(ave_nl = mean(value))

resp_summ1 %>%
  ggplot() +
  geom_point(aes(composition, ave_nl)) +
  facet_wrap( ~ Species)

#### Substrate nonlinearity

```{r}
focus_on <- "Substrate"
ave_nl_substrate <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(composition) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_substrate %>%
  ggplot() +
  geom_point(aes(composition, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

#### Organism nonlinearity

```{r}
focus_on <- "Organism"
ave_nl_organism <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(composition) %>%
  summarise(ave_ave_nl = sum(ave_nl)/unique(num_func_groups))
ave_nl_organism %>%
  ggplot() +
  geom_point(aes(composition, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

#### System nonlinearity

```{r}
focus_on <- "System"
ave_nl_system <- resp_summ1 %>%
  group_by(composition) %>%
  summarise(ave_ave_nl = sum(ave_nl) / (unique(num_func_groups)+4))
ave_nl_system %>%
  ggplot() +
  geom_point(aes(composition, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
ggsave("figures/system_nonlinearity.png", width = 3, height = 2)

### Hysteresis

#### Substrate hysteresis

```{r}
focus_on <- "Substrate"
resp_summ3 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2"),
         var_type == focus_on) %>%
  group_by(composition, num_func_groups, resp_measure, var_type) %>%
  summarise(ave_hyst = mean(value))

resp_summ3 %>%
  ggplot() +
  geom_point(aes(x = composition, y = ave_hyst)) +
  facet_wrap( ~ resp_measure, scales="free") +
  ggtitle(paste("Hysteresis in", focus_on, "response."))

#### Organism hysteresis

```{r}
focus_on <- "Organism"
resp_summ3 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2"),
         var_type == focus_on) %>%
  group_by(composition, num_func_groups, resp_measure, var_type) %>%
  summarise(ave_hyst = sum(value) / unique(num_func_groups))

resp_summ3 %>%
  ggplot() +
  geom_point(aes(x = composition, y = ave_hyst)) +
  facet_wrap( ~ resp_measure, scales="free") +
  ggtitle(paste("Hysteresis in", focus_on, "response."))

#### System hysteresis (separate)

```{r}
resp_summ4 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2")) %>%
  group_by(resp_measure, composition) %>%
  summarise(mean_hyst = sum(value) / (unique(num_func_groups)+4))

resp_summ4 %>%
  ggplot() +
  geom_point(aes(x = composition, y = mean_hyst)) +
  facet_wrap( ~ resp_measure) +
  ggtitle("Hysteresis in system response.")

#### System hysteresis (combined)

```{r}
resp_summ5 <- resp_summ4 %>%
  group_by(composition) %>%
  summarise(mean_mean_hyst = mean(mean_hyst))

resp_summ5 %>%
  ggplot() +
  geom_point(aes(x = composition, y = mean_mean_hyst)) +
  xlab("Functional groups present") +
  ylab("Amount of hysteresis") +
  ylim(0,2) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("figures/composition_system_hysteresis.png", width = 2.5, height = 2.5)

# Results: Varying strength of inhibition

## Introduction

Three parameters set the strength of inhibition (half-inhibition constants):

* h_SR_CB = 300,
* h_O_PB = 100,
* h_O_SB = 100,

These work by the inhibition function:

$\frac{1}{1 + \frac{X}{h_X}}$

```{r}
inhibition
```

Plot some examples:

```{r}
default_parameter_values["h_SR_CB"]
SR <- 10^seq(1.4, 2.5, length = 100)
ggplot() +
  geom_line(aes(x = SR, y = inhibition(SR, default_parameter_values["h_SR_CB"]))) +
  ylab("Multiplicative effect\non CB growth rate")
```

```{r}
default_parameter_values["h_O_PB"]
O <- 10^seq(-0.2, 2.5, length = 100)
ggplot() +
  geom_line(aes(x = O, y = inhibition(O, default_parameter_values["h_O_PB"]))) +
  ylab("Multiplicative effect\non PB growth rate")
```

```{r}
default_parameter_values["h_O_SB"]
O <- 10^seq(-0.2, 2.5, length = 100)
ggplot() +
  geom_line(aes(x = O, y = inhibition(O, default_parameter_values["h_O_SB"]))) +
  ylab("Multiplicative effect\non SB growth rate")
```

```{r}
proportion_change <- 2
default_parameter_values["h_O_SB"]
O <- 10^seq(-0.2, 2.5, length = 100)
ggplot() +
  geom_line(aes(x = O, y = inhibition(O, proportion_change*default_parameter_values["h_O_SB"]))) +
  ylab("Multiplicative effect\non SB growth rate")
```

## Experiment

Increasing the inhibition half saturation constants weakens the inhibition.

```{r}
ssfind_minimum_abundances <- default_minimum_abundances
ssfind_simulation_duration <- 10000
ssfind_simulation_sampling_interval <- ssfind_simulation_duration
ssfind_event_interval <- ssfind_simulation_duration
ssfind_parameters <- default_parameter_values
ssfind_init_state <- c(N_CB = NA, #this one is varied
           N_PB = NA,
           N_SB = NA,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)

grid_num_mult <- 20
grid_num_N <- 2
grid_num_a <- 100

a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
inhib_const_mults <- 2^seq(-1, 0.4, length=grid_num_mult)

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 1e5
initial_SBs <- 1e5

expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    inhib_const_mult = inhib_const_mults,
                    a_O = a_Os)
```

```{r eval = run_sims}
ss_h1 <- ss_by_a_h_N(expt)

saveRDS(ss_h1, "data/ss_h1.RDS")
```

```{r}
ss_h1 <- readRDS("data/ss_h1.RDS")
```

## Results

### Close to default inhibition

```{r}
inhib_const_mults
focal_inhibition_multiplier <- inhib_const_mults[14]

ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

### Double strength inhibition

```{r}
## YES! 0.5 does mean twice as strong inhibition
focal_inhibition_multiplier <- 0.5

ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

### Weaker inhibition

```{r}
## YES! 2 does mean half as strong inhibition
focal_inhibition_multiplier <- 2^0.4

ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

### Nonlinearity

```{r}
x_long <- ss_h1 %>%
  mutate(direction = ifelse(initial_N_CB==1, "up", "down")) %>%
  gather(key = Species, value = Quantity, 3:9) 
x_wide <- x_long %>%
  select(-initial_N_CB) %>%
  spread(key = direction, value=Quantity, drop=T)

resp_summary <- x_wide %>%
  group_by(inhib_const_mult, Species) %>%
  summarise(hyst_1 = mean(abs(log10(up) - log10(down))),
            hyst_2 = get_hysteresis_range(up, down, a),
            nl_up = get_nonlinearity(a, log10(up)),
            nl_down = get_nonlinearity(a, log10(down))) %>%
  ungroup() %>%
  gather(key = resp_measure, value = value, 3:6) %>%
  mutate(
         var_type = ifelse(str_sub(Species, 1, 1)=="N", "Organism", "Substrate")
         )
```

#### All species nonlinearity

```{r}
resp_summ1 <- resp_summary %>%
  filter(resp_measure %in% c("nl_down", "nl_up")) %>%
  group_by(inhib_const_mult, Species, var_type) %>%
  summarise(ave_nl = mean(value))

resp_summ1 %>%
  ggplot() +
  geom_point(aes(inhib_const_mult, ave_nl)) +
  facet_wrap( ~ Species)

#### Substrate nonlinearity

```{r}
focus_on <- "Substrate"
ave_nl_substrate <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(inhib_const_mult) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_substrate %>%
  ggplot() +
  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

#### Organism nonlinearity

```{r}
focus_on <- "Organism"
ave_nl_organism <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(inhib_const_mult) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_organism %>%
  ggplot() +
  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

#### System nonlinearity

```{r}
focus_on <- "System"
ave_nl_system <- resp_summ1 %>%
  group_by(inhib_const_mult) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_system %>%
  ggplot() +
  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
ggsave("figures/system_nonlinearity.png", width = 3, height = 2)

### Hysteresis

#### Substrate hysteresis

>>>>

---
title: "EWS analysis"
author: "Owen Petchey & Jennifer Probst"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 4
    number_sections: true
    code_folding: hide
    toc_float:
      collapsed: true
      smooth_scroll: true
---
